---
title: "100xVolBio_Draft"
author: "Allison Adams"
date: "6/16/2022"
output: html_document
---
This is using a new raw data file, 100xRaw_R_Updated_06_16.xlsx, since I deleted several organism entries, see note in excel file. 

This contains both the volume and the biomass calculations
```{r}
library(tidyverse)
library(writexl)
getwd()
```
Load the excel data file
```{r}
volbio100_draft <- read_csv("100xRaw_R_Updated_06_16.csv")
```
Convert to long format

```{r}
volbio_long <- pivot_longer(data = volbio100_draft,
                             cols = c(8:67),
                             names_to = "sample",
                             values_to = "counts")
```

Rearrange the column orders using the select function, and then change the NAs to 0s using the mutate function.
```{r}
volbio_final <- mutate(volbio_long, counts = ifelse(is.na(counts), 0, counts)) %>% 
  select(sample, Organism, Group, name, shp, sa, la, wi, counts)
```

Volume Equation Calculations: change the cone to cone1 so R won't run all the cones2 and cones

```{r}
volbio_fin_cn <- mutate(volbio_final, shp = ifelse(shp == "cone", "cone1", shp))
```
Run the volume function
```{r}
vol_func <- function(diameter, height, width, shape, counts) {
  pi= 3.1415
  volume = 0
   if(grepl("cone1", shape, ignore.case = TRUE)){
    volume = (pi/12) * diameter^2 * height
   }
  if(grepl("cones2", shape, ignore.case = TRUE)){
    volume = 2*(pi/12) * diameter^2 * height
  }
   if(grepl("sph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^3
   } 
  if(grepl("prosph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(grepl("cyl", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter^2 * height
  }
  if(grepl("ellips", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter * height * width
  }
  if(grepl("recbox", shape, ignore.case = TRUE)){
    volume = diameter * height * width
  }
  if(grepl("prisell", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter * height * width
  }
  if(grepl("prispar", shape, ignore.case = TRUE)){
    volume = 0.5 * diameter * height * width
  }
  if(volume == 0){
    return(NA)
  } else{
    return(volume * counts)
  }
 }
```

Calculate volumes
```{r}
volbio_fin_cn2 <- rowwise(volbio_fin_cn) %>% 
  mutate(volume = vol_func(diameter = sa, height = la, width = wi, shape = shp, counts = counts))
```

Add the biomass function
```{r}
biomass_func <- function(Group, volume) {
  biomass = 0
   if(grepl("ciliate", Group, ignore.case = TRUE)){
    biomass = .23 * volume^.984
  }
   if(grepl("tintinnid", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
   }
   if(grepl("chlorophyte", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
   }
   if(grepl("diatom", Group, ignore.case = TRUE)){
    biomass = .117 * volume^.881
  }
   if(grepl("dinoflagellate", Group, ignore.case = TRUE)){
    biomass = .760 * volume^.819
   }
   if(grepl("flagellate", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("ochrophyte", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("unidentified", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("cyanobacteria", Group, ignore.case = TRUE)){
    biomass = .181 * volume
  }
  if(biomass == 0){
    return(NA)
  } else{
    return(biomass)
  }
 }
```

Calculate Biomass  
```{r}
volbio_all <- rowwise(volbio_fin_cn2) %>% 
  mutate(biomass = biomass_func(volume = volume, Group = Group))
```

Eliminate all the rows that have zero occurrences
```{r}
volbio_all_no0 <- volbio_all %>% 
    rowwise() %>% 
    filter(counts != 0)
```

Create the file as an excel document
write_xlsx(outAll, "add filepath and name here.xlsx")
```{r}
#write_xlsx(volbio_all_no0, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/100x June 2022/VolBioAllNo0.xlsx")
```

Look at the data to see how I can group by size class. Try pivot_wider to see the distribution of organisms by station/sample
Note: after looking at volbio_wide, this isn't as helpful as I expected for seeing organisms by size.

```{r}
volbio_wide <- pivot_wider(volbio_all_no0, names_from = Organism, values_from = biomass)
```

```{r}
#write_xlsx(volbio_wide, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/100x June 2022/VolBioWide.xlsx")
```



