---
title: "100xVolume"
author: "Allison Adams"
date: "6/10/2022"
output: html_document
---
6/10/22
This file is with RawCounts 100x, starting with calculating the volumes so I can then calculate biomass


```{r}
library(tidyverse)
library(writexl)
getwd()
```
Add this at the bottom
install.packages("writexl")
library(writexl)
write_xlsx(outAll, "add filepath and name here.xlsx")

Load the excel data file
```{r}
raw_vol100 <- read_csv("100xRaw_R.csv")
```
Convert to long format

raw_vol_long <- pivot_longer(data = raw_vol100,
                             cols = c("SJR2_site", "SJR2_FC1", 
                                      "SJR2_T241", "SJR2_FC2", 
                                      "SJR2_T242", "SJR2_FC3", 
                                      "SJR2_T243", "SJR2_IC1", 
                                      "SJR2_IC2", "SJR2_IC3", 
                                      "YBP1_site", "YBP1_FC1", 
                                      "YBP1_T241", "YBP1_FC2", 
                                      "YBP1_T242", "YBP1_FC3", 
                                      "YBP1_T243", "YBP1_IC1", 
                                      "YBP1_IC2", "YBP1_IC3", 
                                      "LSZ2_site", "LSZ2_FC1", 
                                      "LSZ2_T241", "LSZ2_FC2", 
                                      "LSZ2_T242", "LSZ2_FC3", 
                                      "LSZ2_T243", "LSZ2_IC1", 
                                      "LSZ2_IC2", "LSZ2_IC3",
                                      "WLD2_site", "WLD2_FC1", 
                                      "WLD2_T241", "WLD2_FC2", 
                                      "WLD2_T242", "WLD2_FC3", 
                                      "WLD2_T243", "WLD2_IC1", 
                                      "WLD2_IC2", "WLD2_IC3",
                                      "SJR1_site", "SJR1_FC1", 
                                      "SJR1_T241", "SJR1_FC2", 
                                      "SJR1_T242", "SJR1_FC3", 
                                      "SJR1_T243", "SJR1_IC1", 
                                      "SJR1_IC2", "SJR1_IC3",),
                             names_to = "sample",
                             values_to = "counts")

In the above code, I forgot to add the YBP2 stations, so it kept those columns.

Also, below is an experiment following Luka's advice to add the cols = for the sample names without writing them all out like I did above. It worked!

```{r}
raw_vol_long <- pivot_longer(data = raw_vol100,
                             cols = c(8:67),
                             names_to = "sample",
                             values_to = "counts")
```

Below is another way to do the above, per Luka. Since all the station columns contain an underscore, I can try to select those columns that contain one using the contains() function
```{r}
raw_vol_long_exp2 <- pivot_longer(data = raw_vol100,
                             cols = contains("_"),
                             names_to = "sample",
                             values_to = "counts")
```
Rearrange the column orders using the select function, and then change the NAs to 0s using the mutate function.
```{r}
raw_vol_final <- mutate(raw_vol_long, counts = ifelse(is.na(counts), 0, counts)) %>% 
  select(sample, Organism, Group, name, shp, sa, la, wi, counts)
```
Make another column for high growth rate and low growth rate stations and reorder the rows according to highs and lows. Use grepl to get R to designate all samples under on station to the respective high or low, as in YBP2_site, YBP2_FC1, and so on.

Luka said, grepl("YBP2", sample)
instead of
sample == "YBP2"

```{r}
raw_vol_sta_fin <- mutate(raw_vol_final, 
   growth_rate = case_when( grepl("YBP2", sample) ~ "high",
                            grepl("SJR1", sample) ~ "high",
                            grepl("LSZ2", sample) ~ "high",
                            grepl("YBP1", sample) ~ "low",
                            grepl("SJR2", sample) ~ "low",
                            grepl("WLD2", sample) ~ "low")) %>% 
  relocate(growth_rate) %>% 
  arrange(growth_rate, sample)
```


This is the chunk of code I ran previously, saving it just in case, but don't want to run it.
raw_vol_sta_fin <- mutate(raw_vol_final, 
   growth_rate = case_when(sample == "YBP2" ~ "high",
                           sample == "SJR1" ~ "high",
                           sample == "LSZ2" ~ "high",
                           sample == "YBP1" ~ "low",
                           sample == "SJR2" ~ "low",
                           sample == "WLD2" ~ "low")) %>% 
  relocate(growth_rate) %>% 
  arrange(growth_rate, sample)
  
This below is from when Wim was trying to help me. He suggested I use ifelse and substring, but since Luka told me how to use grepl, there was no need for me to figure this out.
raw_vol_sta_fin$growth_rate <- 
 ifelse(substring (sample == "YBP2" , "high",
        "SJR1" , "high",)
        
 #### Here I'll try to use the toy code to filter by counts >0:
 cil_ct_by_site_no0 <- group_by(cil_ct_by_site, Organism) %>% 
  filter(sum(counts)>0) %>% 
  ungroup()
cil_ct_by_site_no0  

```{r}
raw_vol_no0 <- group_by(raw_vol_final, name) %>% 
  filter(sum(counts)>0) %>% 
  ungroup()
raw_vol_no0
```
Save raw_vol_no0 as xlsx
```{r}
write_xlsx(raw_vol_no0, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/100x June 2022/Raw_Vol.xlsx")
```

Add the equations function

```{r}
vol_func <- function(diameter, height, width, shape, counts) {
  pi= 3.1415
  volume = 0
   if(grepl("cone", shape, ignore.case = TRUE)){
    volume = (pi/12) * diameter^2 * height
   }
  if(grepl("cones2", shape, ignore.case = TRUE)){
    volume = 2*(pi/12) * diameter^2 * height
  }
   if(grepl("sph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^3
   } 
  if(grepl("prosph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(grepl("cyl", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter^2 * height
  }
  if(grepl("ellips", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter * height * width
  }
  if(grepl("recbox", shape, ignore.case = TRUE)){
    volume = diameter * height * width
  }
  if(grepl("prisell", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter * height * width
  }
  if(grepl("prispar", shape, ignore.case = TRUE)){
    volume = 0.5 * diameter * height * width
  }
  if(grepl("ellipsoid + 2 cones + cylinder", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(volume == 0){
    return(NA)
  } else{
    return(volume * counts)
  }
 }
```

Run the data with the volume equations
```{r}
vol_all <- rowwise(raw_vol_no0) %>% 
  mutate(volume = vol_func(diameter = sa, height = la, width = wi, shape = shp, counts = counts))
```
Not sure what the problem is here. Come back tomorrow

