---
title: "Volume Equations"
author: "Allison Adams"
date: "6/16/2022"
output: html_document
---
## Writing equations for volume calculations: see original file "Equations" 4/7/22
## New File all the way at the bottom that includes new equations:
Cone, Cylinder, Two Cones, Ellipsoid, Prism on Elliptic Base, Prism on Parallelogram, Prolate Spheroid, Rectangular Box, Sphere
## Also note ine 100xRaw_Volume2.Rmd, that I had to change the shape codes for cone, and ellips since they were seen as duplicates.
For the orignal volume equation function file, go to Deskton/New Growth Rate/VolumeEquations2_R.Rmd
```{r}
library(tidyverse)
```

```{r}
vol_func <- function(diameter, height, width, shape, counts) {
  pi= 3.1415
  volume = 0
   if(grepl("cone1", shape, ignore.case = TRUE)){
    volume = (pi/12) * diameter^2 * height
   }
  if(grepl("cones2", shape, ignore.case = TRUE)){
    volume = 2*(pi/12) * diameter^2 * height
  }
   if(grepl("sph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^3
   } 
  if(grepl("prosph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(grepl("cyl", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter^2 * height
  }
  if(grepl("ellips1", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter * height * width
  }
  if(grepl("recbox", shape, ignore.case = TRUE)){
    volume = diameter * height * width
  }
  if(grepl("prisell", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter * height * width
  }
  if(grepl("prispar", shape, ignore.case = TRUE)){
    volume = 0.5 * diameter * height * width
  }
  if(grepl("ellipsoid + 2 cones + cylinder", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(volume == 0){
    return(NA)
  } else{
    return(volume * counts)
  }
 }