---
title: "400x_VolBio"
author: "Allison Adams"
date: "6/23/2022"
output: html_document
---

This file is for calculating volume and biomass of the 400x raw counts and includes code for creating files filtered by individual organism groups, sucha as chlorophytes, ciliates, diatoms, etc.

See 400x_VolBio2.R for the file that has only the volume and biomass code.

This contains both the volume and the biomass calculations
```{r}
library(tidyverse)
library(writexl)
getwd()
```
Load the excel csv data file
```{r}
raw400 <- read_csv("400x_RawCount_R.csv")
```
Convert to long format

```{r}
raw400_long <- pivot_longer(data = raw400,
                             cols = c(9:68),
                             names_to = "sample",
                             values_to = "counts")
```

Rearrange the column orders to have the station/sample column first, using the select function, and then change the NAs to 0s using the mutate function.
```{r}
raw400_fin <- mutate(raw400_long, counts = ifelse(is.na(counts), 0, counts)) %>% 
  select(sample, Organism, Group, type, name, shp, sa, la, wi, counts)
```

Run the volume function
```{r}
vol_func <- function(diameter, height, width, shape, counts) {
  pi= 3.1415
  volume = 0
   if(grepl("cone1", shape, ignore.case = TRUE)){
    volume = (pi/12) * diameter^2 * height
   }
  if(grepl("cones2", shape, ignore.case = TRUE)){
    volume = 2*(pi/12) * diameter^2 * height
  }
   if(grepl("sph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^3
   } 
  if(grepl("prosph", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter^2 * height
  }
  if(grepl("cyl", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter^2 * height
  }
  if(grepl("ellips", shape, ignore.case = TRUE)){
    volume = (pi/6) * diameter * height * width
  }
  if(grepl("recbox", shape, ignore.case = TRUE)){
    volume = diameter * height * width
  }
  if(grepl("prisell", shape, ignore.case = TRUE)){
    volume = (pi/4) * diameter * height * width
  }
  if(grepl("prispar", shape, ignore.case = TRUE)){
    volume = 0.5 * diameter * height * width
  }
  if(volume == 0){
    return(NA)
  } else{
    return(volume * counts)
  }
 }
```

Calculate volumes
```{r}
vol400 <- rowwise(raw400_fin) %>% 
  mutate(volume = vol_func(diameter = sa, height = la, width = wi, shape = shp, counts = counts))
```

If you want to add a column that just calculates volume of one organism and not mulitplying it by the counts, see the file 100x_Corrected_06_17 in the 100x June 2022 file.

Add the biomass function
```{r}
biomass_func <- function(Group, volume) {
  biomass = 0
   if(grepl("ciliate", Group, ignore.case = TRUE)){
    biomass = .23 * volume^.984
  }
   if(grepl("tintinnid", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
   }
   if(grepl("chlorophyte", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
   }
   if(grepl("diatom", Group, ignore.case = TRUE)){
    biomass = .117 * volume^.881
  }
   if(grepl("dinoflagellate", Group, ignore.case = TRUE)){
    biomass = .760 * volume^.819
   }
   if(grepl("flagellate", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("ochrophyte", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("unidentified", Group, ignore.case = TRUE)){
    biomass = .216 * volume^.939
  }
   if(grepl("cyanobacteria", Group, ignore.case = TRUE)){
    biomass = .181 * volume
  }
  if(biomass == 0){
    return(NA)
  } else{
    return(biomass)
  }
 }
```

Calculate Biomass  
```{r}
bio400 <- rowwise(vol400) %>% 
  mutate(biomass = biomass_func(volume = volume, Group = Group))
```

Eliminate all the rows that have zero occurrences
```{r}
bio400_no0 <- bio400 %>% 
    rowwise() %>% 
    filter(counts != 0)
```

Create the file as an excel document
write_xlsx(outAll, "add filepath and name here.xlsx")
```{r}
#write_xlsx(bio400_no0, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/Bio400_no0.xlsx")
```

Look at the data to see how I can group by size class.

First get a list of all the groups in 400x.
```{r}
unique(bio400_no0$Group)
```
If you want to do pivot wider, here's the temmplate:
volbio_wide <- pivot_wider(volbio_all_no0, names_from = Organism, values_from = biomass)

This below is for finding out ways to select the data by "type" so I can look at the sizes and choose what size classes to create to group the organsim types by. See the Dimensions Summaries.R file in 100x_Corrected _06_17 for more details and notes, but I'll do the 400x type analyis here instead of making a separate R file.

## Chlorophytes
```{r}
chlor4_dim_table <- filter(bio400_no0, grepl("^Chloro", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```
```{r}
write_xlsx( chlor4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Chlor4_Dim.xlsx")
```

Ciliates
```{r}
cil4_dim_table <- filter(bio400_no0, grepl("^Ciliate", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```
```{r}
write_xlsx( cil4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Cil4_Dim.xlsx")
```

Cyanobacteria
```{r}
cyano4_dim_table <- filter(bio400_no0, grepl("^Cyanobacteria", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```
```{r}
write_xlsx( cyano4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Cyano4_Dim.xlsx")
```

Diatom
```{r}
dia4_dim_table <- filter(bio400_no0, grepl("^Diatom", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```
```{r}
write_xlsx( dia4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Dia4_Dim.xlsx")
```

Flagellate
```{r}
flag4_dim_table <- filter(bio400_no0, grepl("^Flagellate", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```

```{r}
write_xlsx( flag4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Flag4_Dim.xlsx")
```

Unidentified
```{r}
unid4_dim_table <- filter(bio400_no0, grepl("^Unidentified", Organism)) %>% 
  group_by(Group, type, sa, la, wi) %>%
  summarize(dimension_count = n(), total_count = sum(counts), 
            biomass_total = sum(biomass)) %>% 
  ungroup()
```

```{r}
write_xlsx(unid4_dim_table, "/Users/allisonadams/Documents/Thesis/Microplankton/R Work/400x June 2022/400x Size Classes/Unid4_Dim.xlsx")
```




